---
title: 'Web Data Analysis and Prediction'
author: "Gaurav Mahajan"
date: "1/12/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
linkedin profile: linkedin.com/in/mahajang

Background and Objective:
Perform Statistics and build  Prediction models to forecast demand for Walmart Store

Domain: Retail


Load Data
```{r}
library(plyr) #cleaning
library(dplyr) 
library(e1071)
library(caret)
library(mlbench)
library(ggplot2)
library(tidyverse)
library(funModeling) 
library(Hmisc)
library(car) #vif


setwd("C:/")
internet_data<- read.csv("C:/internet_dataset.csv", header=TRUE)

dim(internet_data)
names(internet_data)
str(internet_data)
#View(internet_data)

summary(internet_data)
```


Data insights:
1. There are ~32K records with 8 variables. Of which, 6 are numeric and 2 are categorical.
2. Means of Bounces are 0.7, Time spent in a page is 73 (secs.), Unique page views is 1 and No. of visits is 1


```{r}
#Plots:
library(funModeling)

freq(internet_data$Continent) #freq plots of categorical variable
freq(internet_data$Sourcegroup) #freq plots of categorical variable
freq(internet_data$Bounces) #freq plots of Numerical variable
freq(internet_data$Exits) #freq plots of Numerical variable
#freq(internet_data$Timeinpage) #freq plots of Numerical variable
freq(internet_data$Uniquepageviews) #freq plots of Numerical variable
freq(internet_data$Visits) #freq plots of Numerical variable
freq(internet_data$BouncesNew) #freq plots of Numerical variable

#Count of Visitors by Continent and Sourcegroup:
Continent_Sourcegrp<- ggplot(internet_data %>% count(Continent, Sourcegroup) %>%    # Group by Continent and Sourcegroup, then count number in each group
         mutate(pct=n/sum(n)),              # Calculate percent within each region
       aes(Continent, n, fill=Sourcegroup)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%")),
            position=position_stack(vjust=0.5))
Continent_Sourcegrp

#Pageviews vs. Visits
Pg_vs_Vi<- ggplot(data = internet_data) + 
  geom_point(mapping = aes(x = Uniquepageviews, y = Visits))
Pg_vs_Vi

```
Insights :

1. ~82% of the Total vistors are coming from N.America (62%) and EU (20%)
2. ~59% of the vistors are coming from Source groups - Google and Direct
3. ~93% of the visitors had viewed the page 1 time
4. ~74% of the users had visited 1 time on the site
5. ~37% of 62% of total visitors in N.America  are from Facebook and Direct
6. There seems to strong coorelation between page views and visits as the graph shows linear trend


Correlation:

```{r}
library(corrplot)
internet_data2<- internet_data[,c(1:2,5:8)] #Data with numerical variables 
summary(internet_data2)
cor = cor(internet_data2)
corrplot(cor, method="number", type = "upper", order = "hclust", 
         tl.col = "black")
```

Insights  :
There is a high correlation (>0.8) between Exits vs. Bounces, UniquePageViews and Visits
There is a medium correlation (0.67) between Bounces vs. UniquePageViews
There is a high correlation (0.81) between Visits vs. UniquePageViews

Model (to check impact of Unique page views on Visits):

```{r}
M1<- lm(Visits  ~ Uniquepageviews, internet_data)
summary(M1)

```

Model Inpterpretation : 
1. Model shows with every 1 user Visit the Uniquepage views increase by 1 (as expected)
2. Model shows Adj. R sq. of 0.66 (sub optimal) with significant p value (2.2e-16) and F-stats: 6.3e+04


```{r}
M2<- lm(Exits  ~ ., internet_data)
summary(M2)

```

Model Inpterpretation (Ques.3):
Variables that impact Exits are Bounces,Sourcegroups(primarily Google, tableausoftware, visualisingdata, Facebook), Timeinpage, Uniquepageviews and Visits. 


Factors impacting Exit page - step-wise models:

```{r}
library(memisc)
M1<-lm(formula=Exits~ Bounces,data=internet_data)
M2<-update(M1,.~.+Sourcegroup)
M3<-update(M2,.~.+Timeinpage)
M4<-update(M3,.~.+Uniquepageviews)
M5<-update(M4,.~.+Visits)
mtable(M1,M2,M3,M4,M5)
summary(M5)

#Multi-collinearity check:
vif(M5)


#Check Regression Assumptions on Non-linearility, multi-collinearity and non-independence of errors
# Line  arity Plots
par(mfrow=c(1,2))
plot(M5)

```


Model Inpterpretation:
1. M5 model gives optimal results with lowest no. of variables compare to all other models.
2. Model gives 0.79 adj. R sq. with p-value < 2.2e-16 and F stat: 1.101e+04
3. Key factors impacting exits are Bounces, Sourcegroup(fb, google, tableau, visualising data), Timeinpage and Uniquepageviews
4. vif less than 10 in all variables of the M5 model


Factors impacting Bounce rate:

```{r}
library(memisc)

names(internet_data)
M1_bounce<-lm(formula=Bounces ~Exits,data=internet_data)
M2_bounce<-update(M1_bounce,.~ +Continent)
M3_bounce<-update(M2_bounce,.~.+Sourcegroup)
M4_bounce<-update(M3_bounce,.~.+Timeinpage)
M5_bounce<-update(M4_bounce,.~.+Uniquepageviews)
M6_bounce<-update(M5_bounce,.~.+Visits)

mtable(M1_bounce,M2_bounce,M3_bounce,M4_bounce,M5_bounce,M6_bounce)

summary(M6_bounce)


# Multi-collinearity check
vif(M6_bounce) # vif less than 10 in all variables of the M6_bounce model


#Check Regression Assumptions on Non-linearility, multi-collinearity and non-independence of errors
# Line  arity Plots
par(mfrow=c(1,2))
plot(M6_bounce)


```


Model Inpterpretation:
1. M6_bounce model gives optimal results with lowest no. of variables compare to all other models.
2. Model gives 0.7 adj. R sq. with p-value < 2.2e-16 and F stat: 4732
3. Key factors impacting Bounces are - Sourcegroup(google, tableau, visualising data), Timeinpage and Uniquepageviews
4. vif less than 10 in all variables of the M6_bounce model




